---
title: "The International Politics of Human Rights"
subtitle: POSC 1020  -- Introduction to International Relations
author: Steven V. Miller
institute: Department of Political Science
titlegraphic: /Dropbox/teaching/clemson-academic.png
date: 
fontsize: 10pt
output:
 beamer_presentation:
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-beamer.tex
    latex_engine: xelatex
    dev: cairo_pdf
    fig_caption: false
    slide_level: 3
make149: true
mainfont: "Open Sans"
titlefont: "Titillium Web"
---


```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='cache/')

knitr::opts_chunk$set(
                  fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      }
                  )
```

```{r loadstuff, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)

library(tidyverse)
library(lubridate)
library(stringr)
library(stevemisc)
library(scales)
library(knitr)
library(grid)
library(gridExtra)
library(artyfarty)
library(drlib)

# LHR <- read_csv("~/Dropbox/data/latent-human-rights/HumanRightsProtectionScores_v2.04.csv")
ITT <- foreign::read.dta("~/Dropbox/data/itt/RevisedSAFINAL_11Mar2013.dta") %>% tbl_df()

# LHR <- read_csv("~/Dropbox/teaching/posc1020/human-rights-1/cache/M_2_Full_Data.csv")
LHR <- read_csv("~/Dropbox/data/latent-human-rights/M_2_Full_Data.csv") %>%
  rename_all(tolower) # changing standard model

# Vdem <- read_csv("~/Dropbox/data/v-dem/8/V-Dem-CY+Others-v8.csv")
Vdem <- readRDS("~/Dropbox/data/v-dem/9/V-Dem-CY-Core-v9.rds") %>% tbl_df()

Polity <- readxl::read_excel("~/Dropbox/data/polity/p4v2017.xls")

Vdem %>% 
  filter(year == 2018) %>% 
  select(country_name, country_text_id, year, v2x_polyarchy,
         v2x_civlib, v2x_civlib_sd, v2x_civlib_codehigh, v2x_civlib_codelow, 
         v2x_clphy, v2x_clphy_sd, v2x_clphy_codehigh, v2x_clphy_codelow, 
         v2x_clpol, v2x_clpol_sd, v2x_clpol_codehigh, v2x_clpol_codelow, 
         v2x_clpriv, v2x_clpriv_sd,  v2x_clpriv_codehigh, v2x_clpriv_codelow, 
         v2clrelig, v2clrelig_sd, v2clrelig_codehigh, v2clrelig_codelow, 
         v2clacfree, v2clacfree_sd, v2clacfree_codehigh, v2clacfree_codelow, 
         v2cltort, v2cltort_sd, v2cltort_codehigh, v2cltort_codelow, 
         v2clkill, v2clkill_sd, v2clkill_codehigh, v2clkill_codelow, 
         v2clfmove, v2clfmove_sd, v2clfmove_codehigh, v2clfmove_codelow, 
         v2cltrnslw, v2cltrnslw_sd, v2cltrnslw_codehigh, v2cltrnslw_codelow, 
         v2clrspct, v2clrspct_sd, v2clrspct_codehigh, v2clrspct_codelow, 
         ) %>%
  mutate(ccode = countrycode::countrycode(country_name, "country.name", "cown")) %>%
  mutate(ccode = ifelse(country_name == "Serbia", 345, ccode)) %>%
  rename(Country = country_name,
         iso3c = country_text_id) %>%
  mutate(EU = ifelse(ccode %in% c(200, 205, 210, 211, 212, 220, 225, 230, 235, 255, 290, 305,
                                  310, 316, 317, 325, 338, 344, 349, 350, 352, 355, 360, 366,
                                  367, 368, 375, 380, 390), 1, 0),
         USA = ifelse(iso3c == "USA", 1, 0),
         World = 1) %>%
  select(Country, USA, EU, World, iso3c, ccode, year, everything()) -> Vdemcivillibs2018

#
multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL, title="", 
                      fontsize = 12, fontfamily = "Helvetica") {
  require(grid)
 
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
 
  numPlots = length(plots)
 
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
 
  if (nchar(title)>0){
    layout <- rbind(rep(0, ncol(layout)), layout)
  }
 
  if (numPlots==1) {
    print(plots[[1]])
 
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), 
                          ncol(layout), 
                          heights = if (nchar(title)>0) {unit(c(0.5, rep(5,nrow(layout)-1)), "null")}
                          else {unit(c(rep(5, nrow(layout))), "null")})))
 
    # Make each plot, in the correct location
    if (nchar(title)>0) {
      grid.text(title, 
                vp = viewport(layout.pos.row = 1, layout.pos.col = 1:ncol(layout)),
                gp = gpar(fontsize = fontsize, fontfamily = fontfamily))
    }
 
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
 
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


# Introduction
### Puzzle(s) for Today

*Why do states care about human rights outside their borders? Why have these concerns not been more successful?*


# Human Rights as International Institution
### Human Rights

We use "human rights" as shorthand for various Western ideals.

- Primary unit of concern is individual and not the community.
    - Think: juxtaposition to socialism and "Asian values."
- Freedom is higher priority than other values
- Focus on political and economic rights

Treatment is heavily indebted to John Locke's notions of natural rights.

### Human Rights

Milestone: the Universal Declaration of Human Rights (UDHR)

- Defines a "common standard of achievement for all people."
- Heavily indebted to then-current events of the Holocaust (see also: Genocide Convention)

Two separate conventions followed around 20 years later:

- International Covenant on Civil and Political Rights (ICCPR)
- International Covenant on Economic, Social, and Cultural Rights (ICESCR)

Curiously: the U.S. supported but did not ratify them as treaties.

- Waited to 1992 to ratify the ICCPR as "not self-executing."
- Has still not ratified the ICESCR.

### So Why Is This Controversial?

States are self-interested and will promote rights they respect (and fight ones they don't).

- Consider U.S. lip service to "human rights" as derivative of Cold War.
- Non-Western states also see this as increasing Western influence.
- Interpretations may differ on implementation (certainly on social spending on health care and education).

### Why Do States Violate Human Rights?

There any number of reasons for human rights abuses.

- Lack of state capacity
- External threats
- Sovereignty concerns
- Human rights may enfranchise regime dissidents, threaten power.

No state is a saint on the human rights frontier, whether it's the U.S. or China.

## Patterns of Human Rights Abuses
<!-- ### -->

```{r human-rights-leaders-fariss, eval=F, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
LHR %>%
  filter(year == max(year)) %>%
  rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  select(year, cow, latentmean, latentsd) %>%
  arrange(-latentmean) %>% 
  mutate(Country = countrycode::countrycode(cow, "cown", "country.name")) %>% head(50) %>%
  filter(!is.na(Country)) %>%
  ggplot(.,aes(reorder(Country, latentmean), latentmean)) + theme_steve_web() +
  geom_point(colour = "#619cff", size = 3) +
  coord_flip() +
  #geom_linerange(aes(x = Country, ymin = latentmean-(1.645*latentsd),
  #                     ymax = latentmean+(1.645*latentsd)),
  #                 lwd = 1, position = position_dodge(width = 1/2)) +
  geom_pointrange(aes(x = Country, y = latentmean, 
                      ymin = latentmean-(1.96*latentsd),
                        ymax =  latentmean+(1.96*latentsd)),
                    lwd = 1/2, position = position_dodge(width = 1/2),
                     fill = "white", color="#3288BD", shape=22) +
  xlab("") + ylab("Mean Latent Human Rights Protection in 2015") +
  labs(title = "The World's Leaders in Human Rights Protections are in Europe or Clustered in Oceania",
       subtitle = "Fariss' model is 'dynamic' and builds on previous years and a myriad of measures, which may explain why some of these Oceanian countries rank high but have diffuse intervals.",
       caption = "Data: Fariss' (2019) Latent Human Rights Protection Scores (v. forthcoming)")
  
```

###

```{r usa-human-rights-ranking-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
# LHR %>%  -> lhr2015

LHR %>%
  rename(value = theta_mean, sd = theta_sd) %>%
    filter(year == max(year)) %>%
  mutate(Country = countrycode::countrycode(cow, "cown", "country.name"),
         Country = ifelse(cow == 816, "Vietnam", Country))  %>%
  select(Country, cow, value, sd) %>%
  rename(ccode = cow) %>%
  mutate(USA = ifelse(Country == "United States", 1, 0),
         upr = value + .5*(sd),
         lwr = value - .5*(sd),
         Indicator = "Human Rights Protections (2015)") %>%
  arrange(-value) %>%
  mutate(rank = seq(1,nrow(.))) -> lhr2015

Vdemcivillibs2018 %>%
  select(ccode, v2x_civlib, v2x_civlib_codehigh, v2x_civlib_codelow) %>%
  rename(value = v2x_civlib,
         upr = v2x_civlib_codehigh,
         lwr = v2x_civlib_codelow) %>% 
  filter(!is.na(ccode)) %>%
  mutate(Indicator = "Civil Liberties Index (2018)",
         USA = ifelse(ccode == 2, 1, 0)) %>%
  arrange(-value) %>%
  mutate(rank = seq(1,nrow(.))) %>%
  bind_rows(lhr2015, .) %>%
  mutate(Country2 = countrycode::countrycode(ccode, "cown", "country.name"),
         Country2 = ifelse(ccode == 816, "Vietnam", Country2)) %>% 
  filter(rank < 40) %>%
  ggplot(.,aes(reorder_within(Country2, value, Indicator), value, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
    geom_pointrange(aes(ymin = lwr,
                        ymax =  upr),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  scale_x_reordered() +
  facet_wrap(~Indicator, scales="free") +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") +
  labs(title = "The World's Leaders in Human Rights Protections are in Europe or Clustered in Oceania",
       subtitle = "The U.S. is conspicuous by how low it scores in both data sets, though there are important explanations for this.",
       caption = "Data: Varieties of Democracy (v. 9); Fariss' (2019) Latent Human Rights Protection Scores (v. forthcoming).")

# 
# lhr2015 <- with(density(lhr2015$latentmean), data.frame(x, y))
# 
# ggplot(data = lhr2015, mapping = aes(x = x, y = y)) + theme_steve_web() +
#     geom_line()  +
#     geom_area(mapping = aes(x = ifelse(x> .373-.435 & x< .373+.435 , x, NA)), fill = "#4d8bc9", alpha=0.4) +
#   #geom_vline(xintercept = .373, linetype="solid", color="#4d8bc9") +
#   # geom_vline(xintercept = .373+.435) +
#   #geom_vline(xintercept = .373-.435) +
#   geom_vline(xintercept = .929, linetype="dashed") +
#   #geom_segment(x=.373+.435, y=0, xend=.373+.435, yend=.26, color="black", size=.5) +
#   #geom_segment(x=.373-.435, y=0, xend=.373-.435, yend=.228, color="black", size=.5) +
#   geom_segment(x=.373, y=0, xend=.373, yend=.268, color="#4d8bc9") +
#   xlab("Latent Human Rights Protection in 2015") + ylab("Density") +
#   labs(title = "Where Does the U.S. Rank on Human Rights? Below Average.",
#        subtitle = "The U.S. is a tricky country to measure for human rights abuses for a myriad of reasons. It does well in some areas and deceptively terrible in others and Amnesty International, aggressively monitor the U.S.",
#        caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)
#        Note: solid blue indicates U.S. mean human rights protection score in 2015. Intervals are the standard deviation around the mean.")

```

###

```{r usa-human-rights-ranking-vdem, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

# Vdem %>% 
#   filter(year == 2017) %>%
#   select(country_text_id, year, v2x_polyarchy) -> v2polyarchy2017
# 
# Polity %>% filter(year == 2017) %>%
#   select(scode, year, polity2) %>%
#   rename(country_text_id = scode) %>%
#   left_join(p2017, .)





Vdemcivillibs2018 %>%
  arrange(-v2x_civlib) %>%
  mutate(`Civil Liberties Index` = seq(1, nrow(.))) %>%
  arrange(-v2x_clphy) %>%
  mutate(`Physical Violence Index` = seq(1, nrow(.))) %>%
  arrange(-v2x_clpol) %>%
  mutate(`Political Civil Liberties Index` = seq(1, nrow(.))) %>%
  arrange(-v2x_clpriv) %>%
  mutate(`Private Civil Liberties Index` = seq(1, nrow(.))) %>%
  arrange(-v2clrelig) %>%
  mutate(`Freedom of Religion` = seq(1, nrow(.))) %>%
  arrange(-v2clacfree) %>%
  mutate(`Freedom of Academic/Cultural Expression` = seq(1, nrow(.))) %>%
  arrange(-v2cltort) %>%
  mutate(`Freedom from Torture` = seq(1, nrow(.))) %>%
  arrange(-v2clkill) %>%
  mutate(`Freedom from Political Killings` = seq(1, nrow(.))) %>%
  arrange(-v2clfmove) %>%
  mutate(`Freedom of Foreign Movement` = seq(1, nrow(.))) %>%
  arrange(-v2cltrnslw) %>%
  mutate(`Transparent Laws/Predictable Enforcement` = seq(1, nrow(.))) %>%
  arrange(-v2clrspct) %>%
  mutate(`Impartial Public Administration` = seq(1, nrow(.))) %>%
  filter(iso3c == "USA") %>% 
  select(Country, `Civil Liberties Index` :ncol(.)) %>%
  gather(Category, Rank, 2:ncol(.)) %>%
  select(-Country) %>%
  rename(`**Category**` = Category,
         `**Rank**` = Rank) %>%
  kable(., align=c("l","c"),
        caption = "U.S. Human Rights Ranks Across Various Vdem Metrics, 2018")


#vec_fontface <- ifelse(Vdemcivillibs2018$Country=="United States of America","bold","plain")
# vec_color <- ifelse(Vdemcivillibs2018$Country=="United States of America","red","black")

```

### 

```{r usa-human-rights-ranking-vdem-other-democracies, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  mutate(Category = "Civil Liberties Index") %>% 
  ggplot(.,aes(reorder(Country, v2x_civlib), v2x_civlib, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  facet_wrap(~Category) +
  geom_pointrange(aes(ymin = v2x_civlib_codelow,
                        ymax =  v2x_civlib_codehigh),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") -> civlibp

Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  mutate(Category = "Physical Violence Index") %>% 
  ggplot(.,aes(reorder(Country, v2x_clphy), v2x_clphy, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  facet_wrap(~Category) +
  geom_pointrange(aes(ymin = v2x_clphy_codelow,
                        ymax =  v2x_clphy_codehigh),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") -> physviop


Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  mutate(Category = "Political Civil Liberties Index") %>% 
  ggplot(.,aes(reorder(Country, v2x_clpol), v2x_clpol, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  facet_wrap(~Category) +
  geom_pointrange(aes(ymin = v2x_clpol_codelow,
                        ymax =  v2x_clpol_codehigh),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") -> polcivlibp

Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  mutate(Category = "Private Civil Liberties Index") %>% 
  ggplot(.,aes(reorder(Country, v2x_clpriv), v2x_clpriv, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  facet_wrap(~Category) +
  geom_pointrange(aes(ymin = v2x_clpriv_codelow,
                        ymax =  v2x_clpriv_codehigh),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") -> privcivlibp


# Vdemcivillibs2018 %>%
#   filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
#   select(Country, v2x_civlib, v2x_civlib_codehigh, v2x_civlib_codelow, 
#          v2x_clphy,v2x_clphy_codehigh, v2x_clphy_codelow, 
#          v2x_clpol, v2x_clpol_codehigh, v2x_clpol_codelow, 
#          v2x_clpriv, v2x_clpriv_codehigh, v2x_clpriv_codelow)

Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  select(Country, v2x_civlib,
         v2x_clphy,
         v2x_clpol, 
         v2x_clpriv) %>%
  gather(Indicator, value, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                          "Civil Liberties Index" = "v2x_civlib",
                                         "Physical Violence Index" = "v2x_clphy",
                                         "Political Liberties Index" = "v2x_clpol",
                                         "Private Civil Liberties Index" = "v2x_clpriv")) -> humanrights_democracies

Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  select(Country, v2x_civlib_codehigh,
         v2x_clphy_codehigh,
         v2x_clpol_codehigh, 
         v2x_clpriv_codehigh) %>%
  gather(Indicator, upr, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                          "Civil Liberties Index" = "v2x_civlib_codehigh",
                                         "Physical Violence Index" = "v2x_clphy_codehigh",
                                         "Political Liberties Index" = "v2x_clpol_codehigh",
                                         "Private Civil Liberties Index" = "v2x_clpriv_codehigh")) %>%
  left_join(humanrights_democracies, .) -> humanrights_democracies


Vdemcivillibs2018 %>%
  filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
  select(Country, v2x_civlib_codelow,
         v2x_clphy_codelow,
         v2x_clpol_codelow, 
         v2x_clpriv_codelow) %>%
  gather(Indicator, lwr, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                          "Civil Liberties Index" = "v2x_civlib_codelow",
                                         "Physical Violence Index" = "v2x_clphy_codelow",
                                         "Political Liberties Index" = "v2x_clpol_codelow",
                                         "Private Civil Liberties Index" = "v2x_clpriv_codelow")) %>%
  left_join(humanrights_democracies, .) -> humanrights_democracies

# library(drlib)
humanrights_democracies %>%
  mutate(USA = ifelse(Country == "United States of America", 1, 0)) %>%
  ggplot(.,aes(reorder_within(Country, value, Indicator), value, color=as.factor(USA), shape=as.factor(USA))) + 
  theme_steve_web() +
    geom_pointrange(aes(ymin = lwr,
                        ymax =  upr),
                    lwd = 1/2, position = position_dodge(width = 1/2)) +
  scale_shape_manual(values = c(19, 17)) +
  scale_color_manual(values = pal("retro")) +
  scale_x_reordered() +
  facet_wrap(~Indicator, scales="free_y") +
  xlab("") + ylab("") +
  coord_flip() +
  theme(legend.position="none") +
  labs(title = "U.S. Human Rights Performances Against Select Democracies, 2018",
       subtitle = "The U.S. ranks above the median worldwide on all human rights indices, but ranks below average among peer democracies.",
       caption = "Varieties of Democracy, v. 9.")


# library(gridExtra)
# grid.arrange(civlibp, physviop,polcivlibp, privcivlibp,
#              # vp=viewport(width=0.1, height=0.1),
#              padding = unit(1.5, "line"),
#              top = textGrob("     U.S. Human Rights Performances Against Select Democracies, 2018",
#                             gp= gpar(fontsize = 18, fontface=2,fontfamily="Titillium Web"),
#                             hjust = 0, x=0,
#              ),
#              bottom = textGrob("Data: Varieties of Democracy, v. 9      ",
#     gp = gpar(fontface = 3, fontsize = 9),
#     hjust = 1,
#     x = 1
#   ),
#              layout_matrix = matrix(c(1,2, 3, 4), ncol=2, byrow=TRUE))


```









```{r usa-human-rights-ranking-vdem-dross, eval=F, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

Vdemcivillibs2018 %>%
  group_by(EU, USA) %>%
  # filter(iso3c == "USA") %>%
  # select(v2x_civlib, v2x_civlib_sd)
  summarize(meancivlib = mean(v2x_civlib),
            sdcivlib = sqrt(mean(v2x_civlib_sd*v2x_civlib_sd)),
            civlib_upper = mean(v2x_civlib_codehigh),
            civlib_lower = mean(v2x_civlib_codelow),
            meanclphy = mean(v2x_clphy),
            sdclphy = sqrt(mean(v2x_clphy_sd*v2x_clphy_sd)),
            clphy_upper = mean(v2x_clphy_codehigh),
            clphy_lower = mean(v2x_clphy_codelow),
            meanclpol = mean(v2x_clpol),
            sdclpol = sqrt(mean(v2x_clpol_sd*v2x_clpol_sd)),
            clpol_upper = mean(v2x_clpol_codehigh),
            clpol_lower = mean(v2x_clpol_codelow),
            meanclpriv = mean(v2x_clpriv),
            sdclpriv = sqrt(mean(v2x_clpriv_sd*v2x_clpriv_sd)),
            clpriv_upper = mean(v2x_clpriv_codehigh),
            clpriv_lower = mean(v2x_clpriv_codelow),
            meanclrelig = mean(v2clrelig),
            sdclrelig = sqrt(mean(v2clrelig_sd*v2clrelig_sd)),
            clrelig_upper = mean(v2clrelig_codehigh),
            clrelig_lower = mean(v2clrelig_codelow),
            meanclacfree = mean(v2clacfree),
            sdclacfree = sqrt(mean(v2clacfree_sd*v2clacfree_sd)),
            clacfree_upper = mean(v2clacfree_codehigh),
            clacfree_lower = mean(v2clacfree_codelow),
            meancltort = mean(v2cltort),
            sdcltort = sqrt(mean(v2cltort_sd*v2cltort_sd)),
            cltort_upper = mean(v2cltort_codehigh),
            cltort_lower = mean(v2cltort_codelow),
            meanclkill = mean(v2clkill),
            sdclkill = sqrt(mean(v2clkill_sd*v2clkill_sd)),
            clkill_upper = mean(v2clkill_codehigh),
            clkill_lower = mean(v2clkill_codelow),
            meanclfmove = mean(v2clfmove),
            sdclfmove = sqrt(mean(v2clfmove_sd*v2clfmove_sd)),
            clfmove_upper = mean(v2clfmove_codehigh),
            clfmove_lower = mean(v2clfmove_codelow),
            meancltrnslw = mean(v2cltrnslw),
            sdcltrnslw = sqrt(mean(v2cltrnslw_sd*v2cltrnslw_sd)),
            cltrnslw_upper = mean(v2cltrnslw_codehigh),
            cltrnslw_lower = mean(v2cltrnslw_codelow),
            meanclrspct = mean(v2clrspct),
            sdclrspct = sqrt(mean(v2clrspct_sd*v2clrspct_sd)),
            clrspct_upper = mean(v2clrspct_codehigh),
            clrspct_lower = mean(v2clrspct_codelow),) %>%
  filter(!(EU == 0 & USA == 0)) %>%
  ungroup() %>%
  mutate(Category = c("United States", "European Union")) %>%
  select(Category, everything(), -EU, -USA) -> usa_eu_civlibs2018
  
  
Vdemcivillibs2018 %>%
  # filter(iso3c == "USA") %>%
  # select(v2x_civlib, v2x_civlib_sd)
  summarize(meancivlib = mean(v2x_civlib),
            sdcivlib = sqrt(mean(v2x_civlib_sd*v2x_civlib_sd)),
            civlib_upper = mean(v2x_civlib_codehigh),
            civlib_lower = mean(v2x_civlib_codelow),
            meanclphy = mean(v2x_clphy),
            sdclphy = sqrt(mean(v2x_clphy_sd*v2x_clphy_sd)),
            clphy_upper = mean(v2x_clphy_codehigh),
            clphy_lower = mean(v2x_clphy_codelow),
            meanclpol = mean(v2x_clpol),
            sdclpol = sqrt(mean(v2x_clpol_sd*v2x_clpol_sd)),
            clpol_upper = mean(v2x_clpol_codehigh),
            clpol_lower = mean(v2x_clpol_codelow),
            meanclpriv = mean(v2x_clpriv),
            sdclpriv = sqrt(mean(v2x_clpriv_sd*v2x_clpriv_sd)),
            clpriv_upper = mean(v2x_clpriv_codehigh),
            clpriv_lower = mean(v2x_clpriv_codelow),
            meanclrelig = mean(v2clrelig),
            sdclrelig = sqrt(mean(v2clrelig_sd*v2clrelig_sd)),
            clrelig_upper = mean(v2clrelig_codehigh),
            clrelig_lower = mean(v2clrelig_codelow),
            meanclacfree = mean(v2clacfree),
            sdclacfree = sqrt(mean(v2clacfree_sd*v2clacfree_sd)),
            clacfree_upper = mean(v2clacfree_codehigh),
            clacfree_lower = mean(v2clacfree_codelow),
            meancltort = mean(v2cltort),
            sdcltort = sqrt(mean(v2cltort_sd*v2cltort_sd)),
            cltort_upper = mean(v2cltort_codehigh),
            cltort_lower = mean(v2cltort_codelow),
            meanclkill = mean(v2clkill),
            sdclkill = sqrt(mean(v2clkill_sd*v2clkill_sd)),
            clkill_upper = mean(v2clkill_codehigh),
            clkill_lower = mean(v2clkill_codelow),
            meanclfmove = mean(v2clfmove),
            sdclfmove = sqrt(mean(v2clfmove_sd*v2clfmove_sd)),
            clfmove_upper = mean(v2clfmove_codehigh),
            clfmove_lower = mean(v2clfmove_codelow),
            meancltrnslw = mean(v2cltrnslw),
            sdcltrnslw = sqrt(mean(v2cltrnslw_sd*v2cltrnslw_sd)),
            cltrnslw_upper = mean(v2cltrnslw_codehigh),
            cltrnslw_lower = mean(v2cltrnslw_codelow),
            meanclrspct = mean(v2clrspct),
            sdclrspct = sqrt(mean(v2clrspct_sd*v2clrspct_sd)),
            clrspct_upper = mean(v2clrspct_codehigh),
            clrspct_lower = mean(v2clrspct_codelow),) %>%
  ungroup() %>%
  mutate(Category = c("World Average")) %>%
  select(Category, everything()) -> world_civlibs2018

bind_rows(world_civlibs2018, usa_eu_civlibs2018) %>%
  group_by(Category) %>% data.frame



  group_by(Country, iso3c, year) %>% 
  arrange(-v2x_civlib) %>% 
  head(50) %>%
   ggplot(.,aes(reorder(Country, v2x_civlib), v2x_civlib)) + 
  theme_steve_web() +
  geom_point(colour = "#619cff", size = 3) +
  # coord_flip() +
  geom_pointrange(aes(x = Country, y = v2x_civlib, 
                      ymin = v2x_civlib_codelow,
                        ymax =  v2x_civlib_codehigh),
                    lwd = 1/2, position = position_dodge(width = 1/2),
                     fill = "white", color="#3288BD", shape=22) +
  xlab("") + ylab("Mean Latent Human Rights Protection in 2015") +
  theme(axis.text.x=
          element_text(face=ifelse(levels(Vdemcivillibs2018$Country)=="United States of America","bold","italic"))) +
  coord_flip()

  gather(Category, value, 4:ncol(.)) %>%
  ggplot(.,aes())

```


###

```{r  usa-can-mex-torture-allegations-comparison-itt, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
ITT %>% 
  filter(cowccode == 2 | cowccode == 20 | cowccode == 70) %>%
  #mutate(Actor = NA,
  #       Actor = ifelse(aocpolice == 1, "Police", Actor),
  #       Actor = ifelse(aocprison == 1, ""))
  group_by(yearbegin, cowccode) %>%
  summarize(sum = n()) %>% 
  bind_rows(.,tibble(yearbegin=seq(1996,1999),
                     cowccode = 20,
                     sum = 0)) %>% 
  mutate(Country = countrycode::countrycode(cowccode, "cown", "country.name")) %>%
  ggplot(.,aes(as.factor(yearbegin), sum, color=Country)) + theme_steve_web() +
  geom_bar(aes(fill = Country), position = "dodge", stat="identity", alpha = I(0.8),color = I("black")) +
  xlab("") + ylab("Number of Amnesty International Torture Allegations") +
  geom_text(aes(label=sum, group=Country), color="black",
            position=position_dodge(width=.9), size=3.5, vjust = -.5, family="Open Sans") +
  labs(title = "The U.S. Experiences About Twice the Allegations of Ill-Treatment and Torture in a Given Year Compared to Its Neighbors",
       subtitle = "Conrad and Moore argue that torture is unobservable but Amnesty International allegations are reliable and valid, even for comparing the U.S. to Mexico here.",
       caption = "Data: Conrad and Moore's Ill-Treatment & Torture (ITT) Data Collection Project (v. 2.1)")
  
  
```


###

```{r usa-torture-allegations-by-actor-type, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocpolice) %>%
  summarize(policesums =n()) %>%
  group_by(yearbegin) %>%
  mutate(nobsyear = sum(policesums),
         percpolice = policesums/nobsyear) %>%
  filter(aocpolice == "Yes") %>%
  select(yearbegin, nobsyear, percpolice) -> policesums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocprison) %>%
  summarize(prisonsums =n()) %>%
  group_by(yearbegin) %>%
  mutate(nobsyear = sum(prisonsums),
         percprison = prisonsums/nobsyear) %>%
  filter(aocprison == "Yes") %>%
  select(yearbegin, percprison) -> prisonsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocmilitary) %>%
  summarize(militarysums =n()) %>%
  mutate(nobsyear = sum(militarysums),
         percmilitary = militarysums/nobsyear) %>%
  filter(aocmilitary == "Yes") %>%
  select(yearbegin, percmilitary) -> militarysums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocimmigrationdetention) %>%
  summarize(immigsums =n()) %>%
  mutate(nobsyear = sum(immigsums),
         percimmig = immigsums/nobsyear) %>%
  filter(aocimmigrationdetention == "Yes") %>%
  select(yearbegin, percimmig) -> immigsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocintelligence) %>%
  summarize(intelsums =n())  %>%
  mutate(nobsyear = sum(intelsums),
         percintel = intelsums/nobsyear) %>%
  filter(aocintelligence == "Yes") %>%
  select(yearbegin, percintel) -> intelsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocparamilitary) %>%
  summarize(parasums =n()) %>%
  mutate(nobsyear = sum(parasums),
         percpara = parasums/nobsyear) %>%
  filter(aocparamilitary == "Yes") %>%
  select(yearbegin, percpara) -> parasums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocunst) %>%
  summarize(unstsums =n()) %>%
  mutate(nobsyear = sum(unstsums),
         percunst = unstsums/nobsyear) %>%
  filter(aocunst == "Yes") %>%
  select(yearbegin, percunst) -> unstsums

left_join(policesums, prisonsums) %>%
  left_join(.,militarysums) %>%
  left_join(.,immigsums) %>%
  left_join(.,intelsums) %>%
  left_join(.,parasums) %>%
  left_join(.,unstsums) -> allegdata

allegdata[is.na(allegdata)] <- 0

allegdata %>%
  gather(Category, Value, percpolice:percunst) %>%
  mutate(Category = fct_recode(Category,
                               "Police" = "percpolice",
                               "Prison" = "percprison",
                               "Military" = "percmilitary",
                               "Immigration Services" = "percimmig",
                               "Intelligence Agency" = "percintel",
                               "Paramilitary" = "percpara",
                               "Unstated" = "percunst")) %>%
  filter(Category != "Unstated" & Category != "Immigration Services" & Category != "Paramilitary") %>%
  mutate(labs = paste0(round(Value*100),"%")) %>%
  ggplot(.,aes(as.factor(yearbegin), Value, color=Category)) + theme_steve_web() +
  scale_fill_brewer(palette="Set1") +
  geom_bar(aes(fill = Category), position = "dodge", stat="identity", alpha = I(0.8),color = I("black")) +
  xlab("") + ylab("Percentage of Amnesty International Torture Allegations in a Given Year") +
  geom_text(aes(label=labs, group=Category), color="black",
            position=position_dodge(width=.9), size=3.5, vjust = -.5, family="Open Sans") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Allegations of Ill-Treatment and Torture in the U.S. Used to Focus on Police and Prisons. Now They Focus on the Military.",
       subtitle = "Allegations of torture by the police bottomed out to near zero in the 2000s but accusations against the military became 3/4ths of all allegations after 9/11.",
       caption = "Data: Conrad and Moore's Ill-Treatment & Torture (ITT) Data Collection Project (v. 2.1).
       Note: Categories are not mutually exclusive, so an allegation can count for both the police and prisons, for example.
       Data omit allegations against unstated actors and against immigration services. 31% of all allegations in 1997 were against immigration services and around 10-20% in a given year are against unstated actors.")
    
    

```


###

```{r usa-human-rights-stagnation-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
Vdem %>%
    filter(country_text_id == "USA") %>%
#    filter(year >= 1945) %>% 
  select(year, v2x_civlib, v2x_civlib_codelow, v2x_civlib_codehigh,
         v2x_clphy, v2x_clphy_codehigh, v2x_clphy_codelow,
         v2cltort_codelow, v2cltort_codehigh, v2cltort) -> usa_over_time


Vdem %>%
    filter(country_text_id == "USA") %>%
  select(year, v2x_civlib, v2x_clphy, v2cltort) %>%
  gather(Indicator, value, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                         "Civil Liberties Index" = "v2x_civlib",
                                         "Physical Violence Index" = "v2x_clphy",
                                         "Freedom From Torture" = "v2cltort")) -> usa_over_time


Vdem %>%
  filter(country_text_id == "USA") %>%
  select(year, v2x_civlib_codehigh, v2x_clphy_codehigh, v2cltort_codehigh) %>%
  gather(Indicator, upr, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                         "Civil Liberties Index" = "v2x_civlib_codehigh",
                                         "Physical Violence Index" = "v2x_clphy_codehigh",
                                         "Freedom From Torture" = "v2cltort_codehigh")) %>%
  left_join(usa_over_time, .) -> usa_over_time

Vdem %>%
  filter(country_text_id == "USA") %>%
  select(year, v2x_civlib_codelow, v2x_clphy_codelow, v2cltort_codelow) %>%
  gather(Indicator, lwr, 2:ncol(.)) %>%
  mutate(Indicator = forcats::fct_recode(Indicator,
                                         "Civil Liberties Index" = "v2x_civlib_codelow",
                                         "Physical Violence Index" = "v2x_clphy_codelow",
                                         "Freedom From Torture" = "v2cltort_codelow")) %>%
  left_join(usa_over_time, .) -> usa_over_time


usa_over_time %>%
  filter(Indicator == "Physical Violence Index") %>%
  filter(year >= 1945) %>%
  ggplot(.,aes(year, value,
               ymin = lwr,
               ymax = upr)) +
  theme_steve_web() +
  geom_ribbon(fill="#619cff", alpha = I(0.2), color = "black") + geom_line() +
  xlab("Year") + ylab("Physical Violence Index (Vdem)") +
  annotate("text", x= 1960, y=.94, label="1960s-1970s:\nPost-Civil Rights Movement gains", hjust="left", 
           family="Open Sans", size=3.5)  +
  annotate("text", x= 1985, y=.85, label="1980s-1990s:\nprison abuse scandals,\nCIA torture campaigns,\nMOVE 9 bombing,\nIncreasing politicization of incarceration", hjust="left", family="Open Sans", size=3.5) +
    annotate("text", x= 2002, y=.79, label="2001-Present:\nWar on Terror human rights abuses", 
             hjust="left", family="Open Sans", size=3.5) +
  scale_x_continuous(breaks=seq(1945, 2020, by=5)) +
  labs(title = "The Physical Violence Index of the United States, 1945-2018",
       subtitle = "The U.S. made considerable progress on human rights in the 1960s and 1970s. Stagnation started in the 1980s and sunk after the War on Terror started.",
       caption = "Data: Varieties of Democracy (v. 9). Note: 1980s human rights abuses annotated to explain downward turns in the Fariss (2019) data (and FH/AI reports).")

# usa_over_time %>%
#   ggplot(.,aes(year, v2x_civlib,
#                ymin = v2x_civlib_codelow,
#                ymax = v2x_civlib_codehigh)) +
#   theme_steve_web() +
#   geom_ribbon(fill="#619cff", alpha = I(0.2), color = "black") + geom_line()
# 
# Vdem %>%
#     filter(country_text_id == "USA") %>%
#   #  filter(year >= 1945) %>% 
#   select(year, v2x_civlib, v2x_civlib_codelow, v2x_civlib_codehigh,
#          v2x_clphy, v2x_clphy_codehigh, v2x_clphy_codelow,
#          v2cltort_codelow, v2cltort_codehigh, v2cltort) %>%
#   ggplot(.,aes(year, v2x_clphy,
#                ymin = v2x_clphy_codelow,
#                ymax = v2x_clphy_codehigh)) +
#   theme_steve_web() +
#   geom_ribbon(fill="#619cff", alpha = I(0.2), color = "black") + geom_line()
# 
# Vdemcivillibs2018 %>%
#   filter(EU == 1 | USA == 1 | ccode %in% c(20, 155, 165, 385, 395, 900, 920, 666, 732, 740)) %>%
#   select(Country, v2x_civlib_codehigh,
#          v2x_clphy_codehigh,
#          v2x_clpol_codehigh, 
#          v2x_clpriv_codehigh) %>%
#   gather(Indicator, upr, 2:ncol(.)) %>%
#   mutate(Indicator = forcats::fct_recode(Indicator,
#                                           "Civil Liberties Index" = "v2x_civlib_codehigh",
#                                          "Physical Violence Index" = "v2x_clphy_codehigh",
#                                          "Political Liberties Index" = "v2x_clpol_codehigh",
#                                          "Private Civil Liberties Index" = "v2x_clpriv_codehigh")) %>%
#   left_join(humanrights_democracies, .) -> humanrights_democracies
# 
#   
# 
# 
# 
# LHR %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
#   filter(cow == 2) %>%
#   mutate(Country = ifelse(cow == 2, "United States", NA),
#          lwr = NA) %>%
#   ggplot(.,aes(year, latentmean, 
#                ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
#   theme_steve_web() +
#   geom_ribbon(fill="#619cff", alpha = I(0.2), color = "black") + geom_line() +
#   ylab("Mean Latent Human Rights Protection") + xlab("Year") +
#   annotate("text", x= 1960, y=2, label="1960s-1970s:\nPost-Civil Rights Movement gains", hjust="left", 
#            family="Open Sans", size=3.5) +
#   annotate("text", x= 1985, y=0, label="1980s-1990s:\nprison abuse scandals,\nCIA torture campaigns,\nMOVE 9 bombing,\nIncreasing politicization of incarceration", hjust="left", family="Open Sans", size=3.5) +
#     annotate("text", x= 2002, y=-1, label="2001-Present:\nWar on Terror human rights abuses", hjust="left", family="Open Sans", size=3.5) +
#   scale_x_continuous(breaks=seq(1950, 2015, by=5)) +
#   labs(title = "Human Rights Protection Scores for the U.S., 1949-2015",
#        subtitle = "The U.S. made considerable progress on human rights in the 1960s and 1970s, which turned south in the 1980s, stagnated in the 1990s, and sunk after the War on Terror.",
#        caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")
```



```{r usa-chn-human-rights-comparison-fariss, eval=F, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}



LHR %>% 
  filter(cow == 2 | cow == 710) %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  mutate(Country = ifelse(cow == 2, "United States", "China"),
         lwr = NA) %>%
  ggplot(.,aes(year, latentmean, group = Country, color=Country,
               ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(aes(fill = Country), alpha = I(0.2), color = "black") + geom_line() +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  scale_x_continuous(breaks = seq(1950, 2015, by = 5)) +
  labs(title = "Human Rights Protection Scores for the U.S. and China, 1949-2015",
       subtitle = "We oberve a clear contrast between the two starting in 1965 but observe the United States' downward trend.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")

```




```{r usa-can-mex-human-rights-comparison-fariss, eval=F, echo=F, fig.height=8.5, fig.width=14}

LHR %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  filter(cow == 2 |  cow == 20 | cow == 70) %>%
  mutate(Country = ifelse(cow == 2, "United States", NA),
         Country = ifelse(cow == 20, "Canada", Country),
         Country = ifelse(cow == 70, "Mexico", Country),
         lwr = NA) %>% filter(year >= 1980) %>%
  ggplot(.,aes(year, latentmean, group = Country, color=Country,
               ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(aes(fill = Country), alpha = I(0.2), color = "black") + geom_line() +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  scale_x_continuous(breaks = seq(1980, 2015, by = 5)) +
  labs(title = "Human Rights Protection Scores for the U.S., Canada, and Mexico, 1980-2015",
       subtitle = "Notice the U.S. is not exactly a world leader in human rights protections despite demonstrations to the contrary.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")

```


## Why Sign These Treaties?
###

*If states have tenuous commitments to human rights, why do they sign these treaties?*

### Why Sign Human Rights Treaties?

Several reasons, but peer effects seem most persuasive.

- New democracies sign them to "lock in" commitments to human rights.
- New democracies sign them to be "like" other, more established democracies.
- Autocracies sign them when they want some kind of respect.
- Autocracies will also sign them as quid pro quo for other deals (e.g. trade).

### Why Sign Human Rights Treaties?

There are other reasons as well, largely from perspective of other states.

- Security motivations and a concern for contagion.
- "Moral" concerns and an empathy for fellow humans.
- Domestic pressure from TANs and interest groups

###

```{r human-rights-stagnation-ciri, eval=T, echo=F, fig.height=8.5, fig.width=14, warning=F, message=F}
LHR204 <- read_csv("~/Dropbox/data/latent-human-rights/HumanRightsProtectionScores_v2.04.csv") %>% rename_all(tolower) 

# se=sd(.)/sqrt(n())

LHR204 %>% 
  group_by(year) %>%
  summarize(meanlhr = mean(latentmean),
            sdlhr = mean(latentsd),
            meanciri = mean(additive, na.rm=T),
            sdciri = sd(additive, na.rm=T),
            seciri = sd(additive, na.rm=T)/sqrt(n())) %>%
  filter(year > 1979) %>%
  ggplot(.,aes(year, meanciri,
               ymin = meanciri - 1.96*seciri, ymax = meanciri + 1.96*seciri)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(alpha = I(0.2), color = "black") + geom_line() +
#  geom_ribbon(aes(y=meanciri, ymin = meanciri - sdciri, ymax = meanciri + sdciri), alpha = I(0.2), color = "black") + geom_line(aes(y=meanciri))  +
  ylab("Mean Physical Integrity Index") + xlab("Year") +
  scale_x_continuous(breaks = seq(1980, 2015, by = 5)) +
  labs(title = "Conventional Wisdom Holds Human Rights Records Have Stagnated Over Time",
       subtitle = "CIRI data show no real improvement in human rights records despite concerted efforts from NGOs and IGOs.",
       caption = "Data: CIRI's Additive Physical Integrity Index via  Fariss' Latent Human Rights Protection Scores (v. 2.04)")

```


###

```{r human-rights-improvement-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, warning=F, message=F}
# LHR <- read_csv("~/Dropbox/teaching/posc1020/human-rights-1/M_2_Full_Data.csv")
#LHR <- read_csv("~/Dropbox/data/latent-human-rights/M_2_Full_Data.csv") %>%
#  rename_all(tolower) # changing standard model

LHR %>%
  group_by(year) %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  summarize(meanval = mean(latentmean),
           # sdlhr = mean(latentsd),
            se = sd(latentmean, na.rm=T)/sqrt(n()),
           Category = "Human Rights Protections") -> improving_human_rights
Vdem %>%
  select(year, v2x_clphy) %>%
  group_by(year) %>%
  filter(year > 1945) %>%
  summarize(meanval = mean(v2x_clphy, na.rm=T),
            se = sd(v2x_clphy,na.rm=T)/sqrt(n()),
            Category = "Physical Violence Index (Vdem)") %>%
  bind_rows(improving_human_rights, .) %>%
  ggplot(.,aes(year, meanval,
               ymin = meanval - 1.96*se, ymax = meanval + 1.96*se)) + 
  facet_wrap(~Category, scales = "free_y") +
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(alpha = I(0.4), color = "black", fill="#619cff") + geom_line() +
#  geom_ribbon(aes(y=meanciri, ymin = meanciri - sdciri, ymax = meanciri + sdciri), alpha = I(0.2), color = "black") + geom_line(aes(y=meanciri))  +
  ylab("Mean Human Rights Score (Higher Values = Better Human Rights)") + xlab("Year") +
  scale_x_continuous(breaks = seq(1950, 2015, by = 5)) +
  labs(title = "Human Rights Records Have Actually Improved Over Time",
       subtitle = "Fariss argues that a ''changing standard of accountability'' can explain the difference between this graph and the conventional wisdom.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming), Varieties of Democracy (v. 9).
       Note: One possible interpretation for these trends, in light of conventional wisdom, is how rare mass killings are.
       Mass killings play a large role in Fariss' ''changing standard'' model while political killings are a big component of the physical violence index.")



```

## Enforcement Problems
### Enforcement Problems

Enforcement of human rights is tricky. Generally:

- The benefits from cooperation are small at best
- The costs of enforcement could be exceedingly high.

How do we explain this?

### Enforcement Problems

A few explanations for enforcement problems:

- Interests: states violating laws have little interest in handcuffing themselves.
- Motives for enforcement may be weak.
    - Naming and shaming may provoke target.
    - Reduces bargaining leverage on other issues.
    - Sanctions impose costs on target as well sanctioning state.

### Enforcement Problems

When are states likely to enforce human rights?

- When TANs are active and persistent (see: anti-apartheid movement)
- Actions against human rights abusers serve larger political interests
- Actions conform to norms of sovereignty and noninterference.

## Improving Human Rights
### Will Protection Improve?

Several innovations provide some promise.

- Transitional justice and the problem of amnesty
- Individual petition to supranational courts
- The International Criminal Court (ICC)
- Material inducements

# Conclusion
### Conclusion

Why protect human rights?

- Pressure from TANs and domestic interest groups
- Empathy for common humanity
- Material interests

### Conclusion

But what's the problem?

- Material benefits are small relative to costs of enforcement.
- Violators are generally highly motivated
- Human rights might, at the most, be a secondary concern.