---
title: "The International Politics of Human Rights"
subtitle: POSC 1020  -- Introduction to International Relations
author: Steven V. Miller
institute: Department of Political Science
titlegraphic: /Dropbox/teaching/clemson-academic.png
date: 
fontsize: 10pt
output:
 beamer_presentation:
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-beamer.tex
    latex_engine: xelatex
    dev: cairo_pdf
    fig_caption: false
    slide_level: 3
make149: true
mainfont: "Open Sans"
titlefont: "Titillium Web"
---


```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='cache/')

knitr::opts_chunk$set(
                  fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      }
                  )
```

```{r loadstuff, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)

library(tidyverse)
library(lubridate)
library(stringr)
library(stevemisc)
library(scales)
library(knitr)

LHR <- read_csv("~/Dropbox/data/latent-human-rights/HumanRightsProtectionScores_v2.04.csv")
ITT <- foreign::read.dta("~/Dropbox/data/itt/RevisedSAFINAL_11Mar2013.dta") %>% tbl_df()

LHR <- read_csv("~/Dropbox/teaching/posc1020/human-rights-1/cache/M_2_Full_Data.csv")

Vdem <- read_csv("~/Dropbox/data/v-dem/8/V-Dem-CY+Others-v8.csv")

```


# Introduction
### Puzzle(s) for Today

*Why do states care about human rights outside their borders? Why have these concerns not been more successful?*


# Human Rights as International Institution
### Human Rights

We use "human rights" as shorthand for various Western ideals.

- Primary unit of concern is individual and not the community.
    - Think: juxtaposition to socialism and "Asian values."
- Freedom is higher priority than other values
- Focus on political and economic rights

Treatment is heavily indebted to John Locke's notions of natural rights.

### Human Rights

Milestone: the Universal Declaration of Human Rights (UDHR)

- Defines a "common standard of achievement for all people."
- Heavily indebted to then-current events of the Holocaust (see also: Genocide Convention)

Two separate conventions followed around 20 years later:

- International Covenant on Civil and Political Rights (ICCPR)
- International Covenant on Economic, Social, and Cultural Rights (ICESCR)

Curiously: the U.S. supported but did not ratify them as treaties.

- Waited to 1992 to ratify the ICCPR as "not self-executing."
- Has still not ratified the ICESCR.

### So Why Is This Controversial?

States are self-interested and will promote rights they respect (and fight ones they don't).

- Consider U.S. lip service to "human rights" as derivative of Cold War.
- Non-Western states also see this as increasing Western influence.
- Interpretations may differ on implementation (certainly on social spending on health care and education).

### Why Do States Violate Human Rights?

There any number of reasons for human rights abuses.

- Lack of state capacity
- External threats
- Sovereignty concerns
- Human rights may enfranchise regime dissidents, threaten power.

No state is a saint on the human rights frontier, whether it's the U.S. or China.

## Patterns of Human Rights Abuses
###

```{r human-rights-leaders-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
LHR %>%
  filter(YEAR == max(YEAR)) %>%
  rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  select(YEAR, COW, latentmean, latentsd) %>%
  arrange(-latentmean) %>% 
  mutate(Country = countrycode::countrycode(COW, "cown", "country.name")) %>% head(50) %>%
  filter(!is.na(Country)) %>%
  ggplot(.,aes(reorder(Country, latentmean), latentmean)) + theme_steve_web() +
  geom_point(colour = "#619cff", size = 3) +
  coord_flip() +
  #geom_linerange(aes(x = Country, ymin = latentmean-(1.645*latentsd),
  #                     ymax = latentmean+(1.645*latentsd)),
  #                 lwd = 1, position = position_dodge(width = 1/2)) +
  geom_pointrange(aes(x = Country, y = latentmean, 
                      ymin = latentmean-(1.96*latentsd),
                        ymax =  latentmean+(1.96*latentsd)),
                    lwd = 1/2, position = position_dodge(width = 1/2),
                     fill = "white", color="#3288BD", shape=22) +
  xlab("") + ylab("Mean Latent Human Rights Protection in 2015") +
  labs(title = "The World's Leaders in Human Rights Protections are in Europe or Clustered in Oceania",
       subtitle = "Fariss' model is 'dynamic' and builds on previous years and a myriad of measures, which may explain why some of these Oceanian countries rank high but have diffuse intervals.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")
  
```

###

```{r usa-human-rights-ranking-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
LHR %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
    filter(YEAR == max(YEAR)) -> lhr2013
lhr2013 <- with(density(lhr2013$latentmean), data.frame(x, y))

ggplot(data = lhr2013, mapping = aes(x = x, y = y)) + theme_steve_web() +
    geom_line()  +
    geom_area(mapping = aes(x = ifelse(x> .373-.435 & x< .373+.435 , x, NA)), fill = "#4d8bc9", alpha=0.4) +
  #geom_vline(xintercept = .373, linetype="solid", color="#4d8bc9") +
  # geom_vline(xintercept = .373+.435) +
  #geom_vline(xintercept = .373-.435) +
  geom_vline(xintercept = .929, linetype="dashed") +
  #geom_segment(x=.373+.435, y=0, xend=.373+.435, yend=.26, color="black", size=.5) +
  #geom_segment(x=.373-.435, y=0, xend=.373-.435, yend=.228, color="black", size=.5) +
  geom_segment(x=.373, y=0, xend=.373, yend=.268, color="#4d8bc9") +
  xlab("Latent Human Rights Protection in 2015") + ylab("Density") +
  labs(title = "Where Does the U.S. Rank on Human Rights? Below Average.",
       subtitle = "The U.S. is a tricky country to measure for human rights abuses for a myriad of reasons. It does well in some areas and deceptively terrible in others and Amnesty International, aggressively monitor the U.S.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)
       Note: solid blue indicates U.S. mean human rights protection score in 2015. Intervals are the standard deviation around the mean.")

```

###

```{r  usa-can-mex-torture-allegations-comparison-itt, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
ITT %>% 
  filter(cowccode == 2 | cowccode == 20 | cowccode == 70) %>%
  #mutate(Actor = NA,
  #       Actor = ifelse(aocpolice == 1, "Police", Actor),
  #       Actor = ifelse(aocprison == 1, ""))
  group_by(yearbegin, cowccode) %>%
  summarize(sum = n()) %>% 
  bind_rows(.,tibble(yearbegin=seq(1996,1999),
                     cowccode = 20,
                     sum = 0)) %>% 
  mutate(Country = countrycode::countrycode(cowccode, "cown", "country.name")) %>%
  ggplot(.,aes(as.factor(yearbegin), sum, color=Country)) + theme_steve_web() +
  geom_bar(aes(fill = Country), position = "dodge", stat="identity", alpha = I(0.8),color = I("black")) +
  xlab("") + ylab("Number of Amnesty International Torture Allegations") +
  geom_text(aes(label=sum, group=Country), color="black",
            position=position_dodge(width=.9), size=3.5, vjust = -.5, family="Open Sans") +
  labs(title = "The U.S. Experiences About Twice the Allegations of Ill-Treatment and Torture in a Given Year Compared to Its Neighbors",
       subtitle = "Conrad and Moore's ITT data measure Amnesty International (AI) allegations arguing that torture is unobservable but AI allegations are reliable and valid, even for comparing the U.S. to Mexico here.",
       caption = "Data: Conrad and Moore's Ill-Treatment & Torture (ITT) Data Collection Project (v. 2.1)")
  
  
```


###

```{r usa-torture-allegations-by-actor-type, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocpolice) %>%
  summarize(policesums =n()) %>%
  group_by(yearbegin) %>%
  mutate(nobsyear = sum(policesums),
         percpolice = policesums/nobsyear) %>%
  filter(aocpolice == "Yes") %>%
  select(yearbegin, nobsyear, percpolice) -> policesums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocprison) %>%
  summarize(prisonsums =n()) %>%
  group_by(yearbegin) %>%
  mutate(nobsyear = sum(prisonsums),
         percprison = prisonsums/nobsyear) %>%
  filter(aocprison == "Yes") %>%
  select(yearbegin, percprison) -> prisonsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocmilitary) %>%
  summarize(militarysums =n()) %>%
  mutate(nobsyear = sum(militarysums),
         percmilitary = militarysums/nobsyear) %>%
  filter(aocmilitary == "Yes") %>%
  select(yearbegin, percmilitary) -> militarysums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocimmigrationdetention) %>%
  summarize(immigsums =n()) %>%
  mutate(nobsyear = sum(immigsums),
         percimmig = immigsums/nobsyear) %>%
  filter(aocimmigrationdetention == "Yes") %>%
  select(yearbegin, percimmig) -> immigsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocintelligence) %>%
  summarize(intelsums =n())  %>%
  mutate(nobsyear = sum(intelsums),
         percintel = intelsums/nobsyear) %>%
  filter(aocintelligence == "Yes") %>%
  select(yearbegin, percintel) -> intelsums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocparamilitary) %>%
  summarize(parasums =n()) %>%
  mutate(nobsyear = sum(parasums),
         percpara = parasums/nobsyear) %>%
  filter(aocparamilitary == "Yes") %>%
  select(yearbegin, percpara) -> parasums

ITT %>%
  filter(cowccode == 2) %>%
  group_by(yearbegin, aocunst) %>%
  summarize(unstsums =n()) %>%
  mutate(nobsyear = sum(unstsums),
         percunst = unstsums/nobsyear) %>%
  filter(aocunst == "Yes") %>%
  select(yearbegin, percunst) -> unstsums

left_join(policesums, prisonsums) %>%
  left_join(.,militarysums) %>%
  left_join(.,immigsums) %>%
  left_join(.,intelsums) %>%
  left_join(.,parasums) %>%
  left_join(.,unstsums) -> allegdata

allegdata[is.na(allegdata)] <- 0

allegdata %>%
  gather(Category, Value, percpolice:percunst) %>%
  mutate(Category = fct_recode(Category,
                               "Police" = "percpolice",
                               "Prison" = "percprison",
                               "Military" = "percmilitary",
                               "Immigration Services" = "percimmig",
                               "Intelligence Agency" = "percintel",
                               "Paramilitary" = "percpara",
                               "Unstated" = "percunst")) %>%
  filter(Category != "Unstated" & Category != "Immigration Services" & Category != "Paramilitary") %>%
  mutate(labs = paste0(round(Value*100),"%")) %>%
  ggplot(.,aes(as.factor(yearbegin), Value, color=Category)) + theme_steve_web() +
  scale_fill_brewer(palette="Set1") +
  geom_bar(aes(fill = Category), position = "dodge", stat="identity", alpha = I(0.8),color = I("black")) +
  xlab("") + ylab("Percentage of Amnesty International Torture Allegations in a Given Year") +
  geom_text(aes(label=labs, group=Category), color="black",
            position=position_dodge(width=.9), size=3.5, vjust = -.5, family="Open Sans") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Allegations of Ill-Treatment and Torture in the U.S. Used to Focus on Police and Prisons. Now They Focus on the Military.",
       subtitle = "Allegations of torture by the police bottomed out to near zero in the 2000s but accusations against the military, nearly non-existent in the 1990s, became 3/4ths of all allegations after 9/11.",
       caption = "Data: Conrad and Moore's Ill-Treatment & Torture (ITT) Data Collection Project (v. 2.1).
       Note: Categories are not mutually exclusive, so an allegation can count for both the police and prisons, for example.
       Data omit allegations against unstated actors and against immigration services. 31% of all allegations in 1997 were against immigration services and around 10-20% in a given year are against unstated actors.")
    
    

```


###

```{r usa-human-rights-stagnation-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}
LHR %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  filter(COW == 2) %>%
  mutate(Country = ifelse(COW == 2, "United States", NA),
         lwr = NA) %>%
  ggplot(.,aes(YEAR, latentmean, 
               ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
  theme_steve_web() +
  geom_ribbon(fill="#619cff", alpha = I(0.2), color = "black") + geom_line() +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  annotate("text", x= 1960, y=2, label="1960s-1970s:\nPost-Civil Rights Movement gains", hjust="left", 
           family="Open Sans", size=3.5) +
  annotate("text", x= 1985, y=0, label="1980s-1990s:\nprison abuse scandals,\nCIA torture campaigns,\nMOVE 9 bombing,\nIncreasing politicization of incarceration", hjust="left", family="Open Sans", size=3.5) +
    annotate("text", x= 2002, y=-1, label="2001-Present:\nWar on Terror human rights abuses", hjust="left", family="Open Sans", size=3.5) +
  scale_x_continuous(breaks=seq(1950, 2015, by=5)) +
  labs(title = "Human Rights Protection Scores for the U.S., 1949-2015",
       subtitle = "The U.S. made considerable progress on human rights in the 1960s and 1970s, which turned south in the 1980s, stagnated in the 1990s, and sunk after the War on Terror.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")
```


###

```{r usa-chn-human-rights-comparison-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, message=F, warning=F}

LHR %>% 
  filter(COW == 2 | COW == 710) %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  mutate(Country = ifelse(COW == 2, "United States", "China"),
         lwr = NA) %>%
  ggplot(.,aes(YEAR, latentmean, group = Country, color=Country,
               ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(aes(fill = Country), alpha = I(0.2), color = "black") + geom_line() +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  scale_x_continuous(breaks = seq(1950, 2015, by = 5)) +
  labs(title = "Human Rights Protection Scores for the U.S. and China, 1949-2015",
       subtitle = "We oberve a clear contrast between the two starting in 1965 but observe the United States' downward trend.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")

```



###

```{r usa-can-mex-human-rights-comparison-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14}

LHR %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  filter(COW == 2 |  COW == 20 | COW == 70) %>%
  mutate(Country = ifelse(COW == 2, "United States", NA),
         Country = ifelse(COW == 20, "Canada", Country),
         Country = ifelse(COW == 70, "Mexico", Country),
         lwr = NA) %>% filter(YEAR >= 1980) %>%
  ggplot(.,aes(YEAR, latentmean, group = Country, color=Country,
               ymin = latentmean - 1.96*latentsd, ymax = latentmean + 1.96*latentsd)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(aes(fill = Country), alpha = I(0.2), color = "black") + geom_line() +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  scale_x_continuous(breaks = seq(1980, 2015, by = 5)) +
  labs(title = "Human Rights Protection Scores for the U.S., Canada, and Mexico, 1980-2015",
       subtitle = "Notice the U.S. is not exactly a world leader in human rights protections despite demonstrations to the contrary.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")

```


## Why Sign These Treaties?
###

*If states have tenuous commitments to human rights, why do they sign these treaties?*

### Why Sign Human Rights Treaties?

Several reasons, but peer effects seem most persuasive.

- New democracies sign them to "lock in" commitments to human rights.
- New democracies sign them to be "like" other, more established democracies.
- Autocracies sign them when they want some kind of respect.
- Autocracies will also sign them as quid pro quo for other deals (e.g. trade).

### Why Sign Human Rights Treaties?

There are other reasons as well, largely from perspective of other states.

- Security motivations and a concern for contagion.
- "Moral" concerns and an empathy for fellow humans.
- Domestic pressure from TANs and interest groups

###

```{r human-rights-stagnation-ciri, eval=T, echo=F, fig.height=8.5, fig.width=14, warning=F, message=F}
LHR <- read_csv("~/Dropbox/data/latent-human-rights/HumanRightsProtectionScores_v2.04.csv")

LHR %>%
  group_by(YEAR) %>%
  summarize(meanlhr = mean(latentmean),
            sdlhr = mean(latentsd),
            meanciri = mean(additive, na.rm=T),
            sdciri = sd(additive, na.rm=T)) %>%
  filter(YEAR > 1979) %>%
  ggplot(.,aes(YEAR, meanciri,
               ymin = meanciri - 1*sdciri, ymax = meanciri + 1*sdciri)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(alpha = I(0.2), color = "black") + geom_line() +
#  geom_ribbon(aes(y=meanciri, ymin = meanciri - sdciri, ymax = meanciri + sdciri), alpha = I(0.2), color = "black") + geom_line(aes(y=meanciri))  +
  ylab("Mean Physical Integrity Index") + xlab("Year") +
  scale_x_continuous(breaks = seq(1980, 2015, by = 5)) +
  labs(title = "Conventional Wisdom Holds Human Rights Records Have Stagnated Over Time",
       subtitle = "CIRI data show no real improvement in human rights records despite concerted efforts from NGOs and IGOs.",
       caption = "Data: CIRI's Additive Physical Integrity Index via  Fariss' Latent Human Rights Protection Scores (v. 2.04)")

```


###

```{r human-rights-improvement-fariss, eval=T, echo=F, fig.height=8.5, fig.width=14, warning=F, message=F}
LHR <- read_csv("~/Dropbox/teaching/posc1020/human-rights-1/M_2_Full_Data.csv")

LHR %>%
  group_by(YEAR) %>% rename(latentmean = theta_mean, latentsd = theta_sd) %>%
  summarize(meanlhr = mean(latentmean),
            sdlhr = mean(latentsd)) %>%
  ggplot(.,aes(YEAR, meanlhr,
               ymin = meanlhr - 1*sdlhr, ymax = meanlhr + 1*sdlhr)) + 
  theme_steve_web() + theme(legend.position="bottom") +
  geom_ribbon(alpha = I(0.4), color = "black", fill="#619cff") + geom_line() +
#  geom_ribbon(aes(y=meanciri, ymin = meanciri - sdciri, ymax = meanciri + sdciri), alpha = I(0.2), color = "black") + geom_line(aes(y=meanciri))  +
  ylab("Mean Latent Human Rights Protection") + xlab("Year") +
  scale_x_continuous(breaks = seq(1950, 2015, by = 5)) +
  labs(title = "Human Rights Records Have Actually Improved Over Time",
       subtitle = "Fariss argues that a ''changing standard of accountability'' can explain the difference between this graph and the conventional wisdom.",
       caption = "Data: Fariss' Latent Human Rights Protection Scores (v. forthcoming)")

```

## Enforcement Problems
### Enforcement Problems

Enforcement of human rights is tricky. Generally:

- The benefits from cooperation are small at best
- The costs of enforcement could be exceedingly high.

How do we explain this?

### Enforcement Problems

A few explanations for enforcement problems:

- Interests: states violating laws have little interest in handcuffing themselves.
- Motives for enforcement may be weak.
    - Naming and shaming may provoke target.
    - Reduces bargaining leverage on other issues.
    - Sanctions impose costs on target as well sanctioning state.

### Enforcement Problems

When are states likely to enforce human rights?

- When TANs are active and persistent (see: anti-apartheid movement)
- Actions against human rights abusers serve larger political interests
- Actions conform to norms of sovereignty and noninterference.

## Improving Human Rights
### Will Protection Improve?

Several innovations provide some promise.

- Transitional justice and the problem of amnesty
- Individual petition to supranational courts
- The International Criminal Court (ICC)
- Material inducements

# Conclusion
### Conclusion

Why protect human rights?

- Pressure from TANs and domestic interest groups
- Empathy for common humanity
- Material interests

### Conclusion

But what's the problem?

- Material benefits are small relative to costs of enforcement.
- Violators are generally highly motivated
- Human rights might, at the most, be a secondary concern.