---
title: "The Future of International Politics: The Rise of China and Globalization"
subtitle: POSC 1020  -- Introduction to International Relations
author: Steven V. Miller
institute: Department of Political Science
titlegraphic: /Dropbox/teaching/clemson-academic.png
date: 
output:
 beamer_presentation:
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-beamer.tex
    latex_engine: xelatex
    dev: cairo_pdf
    fig_caption: false
    slide_level: 3
make149: true
mainfont: "Open Sans"
titlefont: "Titillium Web"
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='cache/')

knitr::opts_chunk$set(
                  fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      }
                  )
```

```{r loadstuff, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

library(tidyverse)
library(stevemisc)
library(stevedata)
library(lubridate)
library(stringr)
library(scales)
library(WDI)
library(ggrepel)



MB <- read_csv("~/Dropbox/data/military-balance/2015/mb2015.csv")
SWIID <- read_csv("~/Dropbox/data/swiid/swiid8_2/swiid8_2_summary.csv")

MB$nukes <- 0
MB$nukes[MB$country == "United States"] <- 4760
MB$nukes[MB$country == "United Kingdom"] <- 225
MB$nukes[MB$country == "Russia"] <- 4300
MB$nukes[MB$country == "Pakistan"] <- 120
MB$nukes[MB$country == "Israel"] <- 80
MB$nukes[MB$country == "India"] <- 110
MB$nukes[MB$country == "France"] <- 300
MB$nukes[MB$country == "China"] <- 250

Trade <- read_csv("~/Dropbox/data/cow/trade/Dyadic_COW_4.0.csv")

library(fredr)

library(pwt9)
data(pwt9.0)


```


# Introduction
### Puzzle(s) for Today

*We've learned a lot of international politics' present and past, but what does the future hold?*


# The Future of International Politics
### The Future of International Politics

1. The proliferation of WMDs
2. The future of American global leadership (i.e. "The Rise of China?")
3. Globalization and its discontents


## The ''Rise of China'' and the Future of American Leadership
###

```{r gdp-forecasts-usa-chn, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

usa_chn_gdp_forecasts %>% 
  mutate(gdp = ifelse(is.na(gdp), f_gdp, gdp)) %>%
  mutate(gdp = gdp/1e12,
         f_lo80 = f_lo80/1e12,
         f_hi80 = f_hi80/1e12) %>%
  # filter(year == 2037)
  ggplot(.,aes(year, gdp,  linetype=country, color = country, fill=country)) +
  theme_steve_web() +
  geom_ribbon(aes(ymin = f_lo80, ymax = f_hi80), alpha = 0.4) +
  geom_line(size=1.5) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(breaks = seq(1960, 2050, by = 10)) +
  xlab("") + ylab("Gross Domestic Product (Observed and Forecasted) in Trillions Constant 2010 US$") +
  geom_vline(xintercept = 2038.5,linetype = "dashed") +
  scale_color_brewer(palette = "Set1") +
  annotate("text", x=2038, y = 35, 
           label = "China surpasses the U.S. in GDP\n(2039)", 
           hjust = 1, family = "Open Sans")  +
  labs(title = "China Should Surpass the U.S. in GDP by 2039",
       linetype = "", color = "", fill = "",
       subtitle = "China's yearly growth in economic activity to date is greater than the growth we observe in U.S. economic output even as (reasonable) worries about China's economic trajectory persist.",
       caption = "Data: World Bank national accounts data, and OECD National Accounts data files. Forecast based on last year in World Bank data (2017).\nSee ?usa_chn_gdp_forecasts in {stevedata} for more.")


```

###

```{r military-balance-usa-chn-2015, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

MB %>%
  mutate(`Defense\nBudget` = mround2(defbudg/sum(defbudg, na.rm=T)),
         `Population` = round((pop/sum(as.numeric(pop), na.rm = T))*100, 2),
         `Active\nPersonnel` = round((active/sum(active, na.rm = T))*100, 2),
         `Size of\nArmy\n(Personnel)` = round((army/sum(army, na.rm = T))*100, 2),
         `Size of\nNavy\n(Personnel)` = round((navy/sum(navy, na.rm = T))*100, 2),
         `Size of\nAir Force\n(Personnel)` = round((af/sum(af, na.rm = T))*100, 2),
         `Total Equipment\n(Army)` = round((armyequip/sum(armyequip, na.rm = T))*100, 2),
         `Total Equipment\n(Navy)` = round((navyequip/sum(navyequip, na.rm = T))*100, 2),
         `Total Equipment\n(Air Force)` = round((afequip/sum(afequip,na.rm=T))*100, 2),
         `Military\nSatellites` = round((otherequip/sum(otherequip, na.rm = T))*100, 2),
         `Nuclear\nWarheads` = round((nukes/sum(nukes, na.rm = T))*100, 2)) %>%
  select(country, `Defense\nBudget`:`Nuclear\nWarheads`) %>%
  filter(country == "United States" | country == "China") %>%
  rename(Country = country) %>%
  gather(variable, value, -Country) %>%
  mutate(lab = paste0(value,"%")) %>%
  ggplot(.,aes(variable, value, color = Country)) + theme_steve_web() +
  geom_bar(aes(fill=Country), stat="identity", position = "dodge", alpha=0.8, color = "black") +
  xlab("Military Indicator") + ylab("Percentage of World Total") +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette="Set1") +
  geom_text(aes(label=lab, group=Country), color="black",
            position=position_dodge(width=.9), 
            size=3.75, vjust = -.5, family = "Open Sans") +
  labs(title = "China's Eventual ''Rise'' Still Comes Amid Major Military Disparity with the United States",
       subtitle = "Generally: China beats U.S. on manpower and never quality or equipment. We expects these trends to persist even as China surpasses the U.S. in economic output.",
       caption = "Data: Military Balance (2015) and Federation of American Scientists (for Nukes)
       Army Equipment: Tanks, recon equipment, armored fighting vehicles, artillery
       Navy Equipment: submarines, principal surface combatants, amphibious equipment,
       Air Force Equipment: Combat-capable aircraft")

```

### What's the Problem With China's Rise?

"Rising" states expect to do better in war.

- As a result, it's more likely to threaten force to revise the status quo.
- A "risen" China is less likely to honor the terms of a pre-power transition agreement.

Indeed, commitment problems magnify the war-proneness of power transitions.

- Further, the U.S. and China have more disagreements than previous cases like Germany and Japan after WWII.

### Is There Cause for Optimism?

Optmistic analysts see promise in China's globalization/leadership efforts.

- e.g. participation in WTO, Asian Infrastructure and Investment Bank

Indeed, China seems to have a vested stake in the current economic/political order, disputes with the U.S. notwithstanding.

- However, Chinese behavior still permits major skepticism.

## Globalization and Its Discontents
###

```{r cell-subscriptions-1980-present, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

WDI(country=c("US", "GB", "DE", "CA", "FR"), indicator = "IT.CEL.SETS.P2", start=1980, end=2020) %>%
  rename(cell = "IT.CEL.SETS.P2", Country = country) %>%
  ggplot(.,aes(year, cell,linetype = Country, group=Country, color=Country)) + 
  theme_steve_web() +
  geom_line(size=1.5) +
  scale_color_brewer(palette="Set1") +
  geom_hline(yintercept = 0, linetype="dashed") +
  scale_x_continuous(breaks = seq(1980, 2020, by =4)) +
  xlab("") + ylab("Mobile Cellular Subscriptions (per 100 people)") +
  labs(caption = "Data: International Telecommunication Union, via the World Bank",
       title = "More and More People Are Owning Cell Phones",
       subtitle = "Globalization has made cell phone ownership readily accessible, which has been a major boon to quality of life in Western countries.")

```

###

```{r usa-household-computers-1984-2015, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

usa_computers %>%
  mutate(lab = paste0(value,"%")) %>%
  ggplot(.,aes(as.factor(year),value)) + theme_steve_web() + 
  geom_bar(stat="identity", position = "dodge", alpha=0.8, color = "black", fill="#00BFC4") +
  geom_text(aes(label=lab), vjust = -.5, colour = "black",
            position = position_dodge(.9), family="Open Sans", size=3.1) +
  ylim(0, 100) +
  xlab("") + ylab("Percentage of Households with a Computer") +
  labs(caption = "Data: Current Population Survey Estimates",
       title = "Almost Every Household Has a Computer These Days, Thanks to Globalization",
       subtitle = "Household computers used to be a luxury item before parts could be cheaply made/imported from Asia. In fact, the old Apple Macintosh from 1984 cost more than $6,000 in 2017 USD.")

# https://www.census.gov/content/dam/Census/library/publications/2018/acs/ACS-39.pdf
  

```

###

```{r usa-price-index-tech-goods-1997-2015, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

LTPT %>% filter(category %nin% c("Internet services and electronic information providers",
                                "Photographic equipment and supplies",
                                "Computer software and accessories")) %>%
  ggplot(.,aes(date, value, color=category, group=category, linetype=category)) + 
  theme_steve_web() +
  geom_line(size=1.1) + 
  scale_color_brewer(palette="Set1") +
  # scale_color_manual(values=c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F")) +
  scale_x_date(date_breaks = "2 years",
               breaks = seq(1996, 2016, by = 2),
               date_labels = "%Y") +
  xlab("") + ylab("") +
  labs(title = "The Price Indices For Computers and TVs Have Declined More Than 95% Since December 1997",
       linetype = "", color="",
       subtitle = "Most of the decline happened between 1998 and 2003 following a boom in cheaper parts and products coming from trading partners in Asia.",
       caption = "Data: Bureau of Labor and Statistics. Base: December 1997.\nSource: https://www.bls.gov/opub/ted/2015/long-term-price-trends-for-computers-tvs-and-related-items.htm.\nSee: ?LTPT in {stevedata}")

```

###

```{r let-them-watch-tv, echo=F, eval=T, fig.width = 14, fig.height = 9, warning = F, message = F}
ggplot() +
  geom_line(data = LTWT, aes(date, value, color = category), size=1.1) +
  theme_steve_web() +
  guides(color = FALSE) +
  geom_hline(yintercept = 100, linetype = "dashed") +
  scale_x_date(limits = as.Date(c("2000-01-01","2022-01-01")),
               date_labels = "%Y",
               date_breaks = "2 year") +
  scale_colour_brewer(palette = "RdBu") +
  # scale_x_date(#date_breaks = "1 year",
  #              breaks = seq(as.Date("2000-01-01"),
  #                           as.Date("2022-12-31"), by="1 year"),
  #              #date_minor_breaks = "3 months",
  #              date_labels = "%Y") +
  #  geom_line(size=1.1) +
  geom_text(data = LTWT %>% filter(date == "2019-12-01"), aes(label = category,
                                                                           x = as.Date("2020-01-01"),
                                                                           y = value,
                                                                           hjust = 0,
                                                                           color = category,
                                                                           fontface = "bold",
                                                                           family = "Open Sans"),
            size = 3) +
  xlab("Date") + ylab("Consumer Price Index for Urban Consumers") +
  labs(title = "''Let Them Watch TV''",
       caption = "Data: Bureau of Labor Statistics. Base period = Jan. 1, 2000.",
       subtitle = "Liberalized trade has made consumer electronics (like TVs) fractions of their past prices. Yet, young adults face mounting costs for college, child-raising,\nand health care that government policy has failed to address.")

```

###

```{r usa-rok-trade-balance-1949-2018, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}


Trade %>% filter(ccode1 == 2  & ccode2 == 732) %>%
  mutate(deficit = flow2 - flow1) %>%
  select(year, deficit) %>%
  bind_rows(., tibble(year = c(2015, 2016, 2017, 2018),
                      deficit = c(-28273.2, -27571.8, -22887.4,-17946.4 ))) %>% 
  ggplot(.,aes(year, deficit)) + theme_steve_web() +
  geom_bar(stat="identity", position = "dodge", alpha=0.8, color = "black", fill="#F8766D") +
  scale_x_continuous(breaks=seq(1950, 2015, by = 5)) +
  scale_y_continuous(labels = scales::comma) +
  xlab("Year") + ylab("Balance of Trade, in Millions (Negative Values = ''Deficits'')") +
  labs(title = "The Boom in Cheap Household Computers Followed a Surge of Imports from South Korea (Among Others)",
       subtitle = "South Korea specializes in cheaply producing high-quality consumer electronics for consumption in the U.S.",
       caption = "Data: Correlates of War Dyadic Trade Data (v. 4.0, 1949-2014). U.S. Census Bureau: 2015-2018.")

```

###

```{r usa-manufacturing-employment-1987-2018, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

fredr(series_id = "PRS30006013",
             observation_start = as.Date("1947-01-01")) %>%
  ggplot(.,aes(date, value)) + geom_line(size=1.5, color="#00BFC4") + 
  theme_steve_web() +
  scale_x_date(date_breaks = "2 years",
               breaks = seq(1988, 2018, by = 2),
               date_labels = "%Y",
               limits = as.Date(c('1986-01-01','2019-01-01'))) +
  xlab("") + ylab("Manufacturing Employment (Seasonally Adjusted, Indexed [2012])") +
  labs(caption = "Data:  U.S. Bureau of Labor Statistics. Data are seasonally adjusted with 2012 as the index year. Shaded areas are recessions.",
       title = "U.S. Manufacturing Employment Has Declined as Globalization Has Increased",
       subtitle = "You'll almost always hear of this as ''jobs leaving the country'' because of globalization.") +
  annotate("rect", xmin=as.Date("1990-07-01"), xmax=as.Date("1991-04-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2001-04-01"), xmax=as.Date("2001-10-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2008-01-01"), xmax=as.Date("2009-07-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  ylim(90, 155)

```

###

```{r usa-manufacturing-output-1987-2018, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

fredr(series_id = "OUTMS",
             observation_start = as.Date("1987-01-01")) %>%
  ggplot(.,aes(date, value)) + geom_line(size=1.5, color="#00BFC4") + theme_steve_web() +
  scale_x_date(date_breaks = "2 years",
               breaks = seq(1986, 2018, by = 2),
               date_labels = "%Y",
               limits = as.Date(c('1986-01-01','2019-01-01'))) +
  xlab("") + ylab("Manufacturing Output (Seasonally Adjusted, Indexed [2012])") +
  labs(caption = "Data:  U.S. Bureau of Labor Statistics. Data are seasonally adjusted with 2012 as the index year. Shaded areas are recessions.",
       title = "U.S. Manufacturing Output Has Generally Risen Despite the Drop in Manufacturing Employment",
       subtitle = "Automation explains more of this phenomenon than globalization, but globalization gets the blame.") +
  annotate("rect", xmin=as.Date("1990-07-01"), xmax=as.Date("1991-04-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2001-04-01"), xmax=as.Date("2001-10-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2008-01-01"), xmax=as.Date("2009-07-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  ylim(60, 140)

```

###


```{r usa-manufacturing-compensation-1987-2018, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}
fredr(series_id = "COMPRMS",
             observation_start = as.Date("1987-01-01")) %>%
  ggplot(.,aes(date, value)) + geom_line(size=1.5, color="#00BFC4") + theme_steve_web() +
  scale_x_date(date_breaks = "2 years",
               breaks = seq(1986, 2018, by = 2),
               date_labels = "%Y",
               limits = as.Date(c('1986-01-01','2019-01-01'))) +
  xlab("") + ylab("Manufacturing Compensation per Hour (Seasonally Adjusted, Indexed [2012])") +
  labs(caption = "Data:  U.S. Bureau of Labor Statistics. Data are seasonally adjusted with 2012 as the index year. Shaded areas are recessions.",
       title = "Manufacturing Pays More Now Than It Did When There Was More Manufacturing Employment",
       subtitle = "Real compensation per hour is on the rise because automation/globalization has decreased demand for low-skilled labor and increased demand for high-skilled labor. This raises wages.") +
  annotate("rect", xmin=as.Date("1990-07-01"), xmax=as.Date("1991-04-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2001-04-01"), xmax=as.Date("2001-10-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2008-01-01"), xmax=as.Date("2009-07-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66")
```


###

```{r usa-labor-share-income-1987-2018, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}
fredr(series_id = "PRS30006173",
             observation_start = as.Date("1987-01-01")) %>%
  ggplot(.,aes(date, value)) + geom_line(size=1.5, color="#00BFC4") + theme_steve_web() +
  scale_x_date(date_breaks = "2 years",
               breaks = seq(1986, 2016, by = 2),
               date_labels = "%Y",
               limits = as.Date(c('1986-01-01','2016-01-01'))) +
  xlab("") + ylab("Manufacturing Compensation per Hours (Seasonally Adjusted, Indexed [2012])") +
  labs(caption = "Data:  U.S. Bureau of Labor Statistics. Data are seasonally adjusted with 2012 as the index year. Shaded areas are recessions.",
       title = "Labor Share of Income in Manufacturing Has Gone Down",
       subtitle = "Gains in income/productivity result in wages for skilled labor/management or reinvestment in capital.") +
  annotate("rect", xmin=as.Date("1990-07-01"), xmax=as.Date("1991-04-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2001-04-01"), xmax=as.Date("2001-10-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66") +
  annotate("rect", xmin=as.Date("2008-01-01"), xmax=as.Date("2009-07-01"), ymin=-Inf, ymax=Inf, alpha = .6, fill = "gray66")
```



###

```{r labor-share-income-1950-2014, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}
pwt9.0 %>% tbl_df() %>%
  mutate(region = countrycode::countrycode(isocode, "iso3c", "region23")) %>%
  mutate(category = ifelse(region %in% c("Northern America",
                                         "Western Europe",
                                         "Australia and New Zealand",
                                         "Northern Europe"),
                           "Advanced Western Economy", "Other Economy")) %>% 
  group_by(year, category) %>%
  summarize(labsh = mean(labsh, na.rm=T))  %>%
  ggplot(.,aes(year, labsh, color=category, linetype=category)) + theme_steve_web() +
  geom_line(size=1.1) + scale_color_brewer(palette="Set1") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(1950, 2015, by = 5)) +
  xlab("") + ylab("Labor Share of Income") +
  labs(caption = "Data: Penn World Table 9.0\n''Advanced Western Economy'' determined by World Bank region for convenience. These observations are the U.S., Canada, Australia, New Zealand, and countries in Western/Northern Europe.",
       linetype = "", color="",
       title = "It's Not Just the United States: Labor Share of Income is Falling Everywhere",
       subtitle = "The trend has been observable as long we have data with more pronounced slides happening in the 1980s.")
```

###

```{r what-explains-declining-share-of-income-imf, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}
tibble::tribble(
  ~Group, ~`Actual change`, ~Technology, ~`GVC Participation`, ~`Financial Integration`, ~`Policy/institutions`, ~Unexplained,
 "AEs",          -3.55,       -1.64,              -0.65,                  -0.48,                -0.17,        -0.61,
  "EMs",          -5.28,       -0.31,              -4.37,                   0.96,                -0.13,        -1.43,
  "EMs exluding China",          -2.58,       -1.24,                 -3,                   1.12,                -0.16,         0.69
  ) %>%
  gather(Category, Value, `Actual change`:Unexplained) %>%
  filter(Category != "Actual change" & Group == "AEs") %>%
  mutate(perc = Value/sum(Value),
         label = paste0(mround2(perc),"%")) %>%
  ggplot(.,aes(Category, perc)) + theme_steve_web() +
  geom_bar(stat="identity", color="black", alpha=0.8) +
  xlab("") + ylab("Percent of Explained Decrease in Labor Share of Income") +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label=label), vjust = -.5, colour = "black",
            position = position_dodge(.9), family="Open Sans") +
  labs(title = "What Explains Declining Share of Income in Advanced Economies? Technology.",
       subtitle = "An IMF report from April 2017 found most  of the movement is explained by declining relative price of investment and initial exposure to routinization across sectors.",
       caption = "Data: IMF. Reconfiguration of Figure 3.11 in ''World Economic Outlook, April 2017: Gaining Momentum?''")


```

###

```{r income-inequality-labor-share-income, echo=F, eval=F, fig.width = 14, fig.height = 8.5, warning = F, message = F}
# I need to work on this...
SWIID %>%
  select(country:gini_disp) %>%
  mutate(ccode = countrycode::countrycode(country, "country.name", "cown"),
         ccode = ifelse(country == "Serbia", 345, ccode)) %>%
  group_by(ccode, year) %>% filter(!is.na(ccode)) %>% 
  summarize(gini_disp = mean(gini_disp)) %>%
  group_by(ccode) %>%
  mutate(gini_diff = gini_disp - lag(gini_disp, 1)) %>%
  ungroup() -> swiid

pwt9.0 %>% tbl_df() %>%
  select(country, year, labsh) %>%
  mutate(ccode = countrycode::countrycode(country, "country.name", "cown"),
        ccode = ifelse(country == "Serbia", 345, ccode)) %>%
  group_by(ccode, year) %>% filter(!is.na(ccode)) %>%
  group_by(ccode) %>%
  mutate(region = countrycode::countrycode(ccode, "cown", "region23")) %>%
  mutate(category = ifelse(region %in% c("Northern America",
                                         "Western Europe",
                                         "Australia and New Zealand",
                                         "Northern Europe"),
                           "Advanced Western Economy", "Other Economy")) %>%
  filter(category == "Advanced Western Economy") %>%
  mutate(labsh_diff = labsh - lag(labsh)) %>%
  left_join(swiid, .) %>% 
  ggplot(.,aes(labsh, gini_disp)) + geom_point() + facet_wrap(~region)


SWIID %>%
  mutate(country = ifelse(country == "United States", "United States of America", country)) %>%
  select(country, year, gini_disp) %>%
  left_join(., pwt9.0) %>%
  select(country, year, gini_disp, labsh) %>%
  mutate(region = countrycode::countrycode(country, "country.name", "region23")) %>%
  mutate(Category = ifelse(region %in% c("Northern America",
                                         "Western Europe",
                                         "Australia and New Zealand",
                                         "Northern Europe"),
                           "Advanced Western Economy", "Other Economy")) %>%
  filter(Category == "Advanced Western Economy" & year > 1970) %>%
  ggplot(.,aes(gini_disp, labsh)) + theme_steve_web() +
  geom_point() + geom_smooth(method="loess") + xlim(20, 36) +
  xlab("GINI") + ylab("Labor Share of Income") +
  labs(title = "Income Inequality Generally Rises in the West as Labor Share of Income Falls",
       subtitle = "Labor share of income decline happens for a variety of reasons but can have important political/economic effects.",
       caption = "Data: Penn World Table (v. 9.0). SWIID (v. 6.2)")

```

<!-- ### -->

```{r income-inequality-labor-share-income-usa, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}
SWIID %>%
  select(country, year, gini_disp) %>%
  filter(country == "United States") %>%
  mutate(country = "United States of America") %>%
  left_join(., pwt9.0) %>%
  select(country, year, gini_disp, labsh) %>%
  filter(country == "United States of America") %>%
  mutate(z_gini = scale(gini_disp),
         z_labsh = scale(labsh)) %>% 
  gather(Category, value, z_gini:z_labsh) %>%
  mutate(Category = ifelse(Category == "z_gini", "GINI", "Labor Share of Income")) %>%
  ggplot(.,aes(year, value, color=Category, linetype=Category)) + theme_steve_web() +
  geom_line(size=1.1) + scale_color_brewer(palette="Set1") +
  scale_x_continuous(breaks = seq(1960, 2016, by = 4)) +
  xlab("") + ylab("GINI or Labor Share of Income (Standardized)") +
  labs(title = "The Correlation Between Income Inequality and Labor Share of Income Is Particularly Strong in the U.S. (r = -.702)",
       subtitle = "Labor share of income's decline is global and has multiple reasons but U.S. policy may be making its effects even worse.",
       caption = "Data: Penn World Table (v. 9.0). SWIID (v. 8.2)")
```

###

```{r income-inequality-usa-ukg-fra-gmy-can, echo=F, eval=T, fig.width = 14, fig.height = 8.5, warning = F, message = F}

SWIID %>%
  filter(country == "United States" | country == "United Kingdom" | country == "France" | country == "Germany" | country == "Canada") %>%
  rename(Country = country) %>%
  ggplot(.,aes(year, gini_disp, color = Country, linetype=Country)) + theme_steve_web() +
  xlab("Year") + ylab("GINI") +
  scale_x_continuous(breaks = seq(1960, 2015, by = 5)) +
  geom_line(size=1.5) +
  scale_color_brewer(palette="Set1") +
  theme(legend.position="bottom") +
  labs(title = "Globalization Skepticism Appears Most Concentrated in Western Countries With Severe Income Inequality",
       subtitle = "Globalization and trade become easy scapegoats to justify austerity measures and cuts to social spending that only compound the problem.",
       caption = "Data: SWIID (v. 6.2)")

```



```{r wvs-swiid-income-inequality-scapegoating-immigrants, echo=F, eval=F, fig.width = 14, fig.height = 8.5, warning = F, message = F}
# Andorra: 28.22657
# South Korea: 31

wvs_immig %>%
  mutate(immig_prohibit = ifelse(e143 > 3, 1, 0)) %>%
  group_by(country, s002, s020) %>%
  summarize(mean_prohibit = mean(immig_prohibit, na.rm=T)) %>%
  filter(s002 == 5) %>%
  mutate(country = ifelse(country == "Great Britain", "United Kingdom", country)) -> wvs_immig_sum5

SWIID %>%
  filter(year >= 2000 & year <= 2008) %>%
  group_by(country) %>%
  summarize(meangini = mean(gini_disp, na.rm=T)) %>%
  left_join(wvs_immig_sum5, .) %>%
  ungroup() %>% 
  mutate(meangini = ifelse(country == "South Korea", 31, meangini)) %>%
  mutate(region = countrycode::countrycode(country, "country.name", "region23")) %>%
  mutate(category = ifelse(region %in% c("Northern America",
                                         "Western Europe",
                                         "Australia and New Zealand",
                                         "Northern Europe", "Southern Europe", "Eastern Europe"),
                           "Advanced Economy", "Other Economy"),
         category = ifelse(country == "Serbia" | country == "Moldova" | country == "Ukraine", "Other Economy", category),
         category = ifelse(country == "Japan" | country == "Chile" | country == "Cyprus", "Advanced Economy", category)) -> swiid_immig


swiid_immig %>%
  ggplot(.,aes(meangini, mean_prohibit)) + theme_steve_web() +
  geom_point(aes(color=category)) + 
  facet_wrap(~category) +
  geom_smooth(method="loess") +
  geom_text_repel(aes(meangini, mean_prohibit, label = country)) +
  xlab("Average GINI (2000-2008)") + ylab("Percent Saying Would-Be Immigrants Should Be Prohibited From Entering Country") +
  labs(title = "Higher Income Inequality Coincides With Societies Scapegoating Immigrants As Part of the Problem",
       subtitle = "Less-developed countries are generally more anti-immigrant but the correlation is near identical in more developed economies.",
       caption = "Data: World Values Survey (Wave 5). SWIID (v. 6.2)")


filter(is.na(meangini))
  summarize(cor = cor(mean, meangini, use="complete.obs")) 



SWIID %>%
  filter(year >= 2000 & year <= 2008) %>%
  group_by(country) %>%
  summarize(meangini = mean(gini_disp, na.rm=T)) %>%
  rename(Country = country) %>%
  left_join(Immig, .) %>%
  mutate(meangini = ifelse(Country == "South Korea", 31, meangini),
         meangini = ifelse(Country == "Andorra", 28.22657, meangini)) %>%
  filter(dummy == 1) %>%
  mutate(region = countrycode::countrycode(Country, "country.name", "region")) %>%
  mutate(Category = ifelse(region %in% c("Northern America",
                                         "Western Europe",
                                         "Australia and New Zealand",
                                         "Northern Europe", "Southern Europe", "Eastern Europe"),
                           "Advanced Economy", "Other Economy"),
         Category = ifelse(Country == "Serbia" | Country == "Moldova" | Country == "Ukraine", "Other Economy", Category),
         Category = ifelse(Country == "Japan" | Country == "Chile" | Country == "Cyprus", "Advanced Economy", Category)) -> giniimmig

giniimmig %>%
  group_by(Category) %>%
  summarize(cor = cor(sum, meangini)) -> cors

giniimmig %>%
  ggplot(.,aes(meangini, sum)) + theme_steve_web() +
  geom_point(aes(color=Category)) + geom_smooth(method="loess") +
  geom_text_repel(aes(meangini, sum, label = Country)) +
  xlab("Average GINI (2000-2008)") + ylab("Percent Saying Would-Be Immigrants Should Be Prohibited From Entering Country") +
  labs(title = "Higher Income Inequality Coincides With Societies Scapegoating Immigrants As Part of the Problem",
       subtitle = "Less-developed countries are generally more anti-immigrant but the correlation is near identical in more developed economies.",
       caption = "Data: World Values Survey (Wave 5). SWIID (v. 6.2)")

```

# Conclusion
### Conclusion

China's "rise" may lead to war with the U.S.

- Preventing this requires integrating China into U.S.-created institutions.

Globalization is a force for peace, but it's a policy that creates winners and losers.

- How you compensate the losers will condition the future of globalization.